\documentclass[10pt, letterpaper]{article}

% --- 宏包加载 ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.5in]{geometry} % 窄页边距确保一页纸
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{setspace}
\doublespace

% --- 样式设置 ---
\definecolor{primary}{RGB}{0, 0, 0}
\urlstyle{same}

\titleformat{\section}{\large\bfseries\uppercase}{}{0pt}{}[\titlerule]
\titlespacing{\section}{0pt}{8pt}{6pt}

\setlist[itemize]{leftmargin=1.5em, nosep, topsep=2pt}

\begin{document}

% --- 个人信息 ---
\begin{center}
    {\Huge \textbf{LI CHEN}} \\
    \vspace{4pt}
    Aspring and Ambitous Astronomy Student 
\end{center}

% --- 教育背景 ---
\section{Education}
\textbf{Gleneagle Secondary School} \hfill Coquitlam, BC, Canada \\
\textit{High School Student} \hfill 2023 -- Present \\
\textbf{Academic Achievements}:
\begin{itemize}
    \item \textbf{Self-Taught AP Curriculum (All Scores of 5)}: Calculus BC, Physics C: Mechanics, Physics 1, Physics 2, Precalculus.
\end{itemize}

% --- 荣誉与奖项 ---
\section{Honors \& Awards}
\begin{itemize}
    \item \textbf{British Physics Olympiad (BPhO) Round 1 | Top Gold} \hfill 2025
    \begin{itemize}
        \item Scored 78/100, ranking in the \textbf{Top 2\%} globally in the UK's premier physics competition.
    \end{itemize}
    \item \textbf{3rd International Olympiad on Astronomy \& Astrophysics - Jr. (IOAA-jr) | Gold Medal} \hfill 2024
    \begin{itemize}
        \item \textbf{Ranked 6th Globally} among 100+ competitors from 19 countries.
    \end{itemize}
    \item \textbf{2nd International Olympiad on Astronomy \& Astrophysics - Jr. (IOAA-jr) | Silver Medal} \hfill 2023
    \begin{itemize}
        \item Ranked 11th globally at age 14 among 80 competitors.
    \end{itemize}
    \item \textbf{Chinese National Astronomy Olympiad (CNAO) | Gold Medal} \hfill 2022
    \begin{itemize}
        \item Youngest medalist of the year (Grade 8) in China's most prestigious astronomy competition.
    \end{itemize}
\end{itemize}

% --- 学术领导力与研究潜力 ---
\section{Academic Leadership \& Research Potential}
\textbf{Physics \& Astronomy Clubs} | \textit{President} \hfill 2024 -- Present
\begin{itemize}
    \item Lead weekly sessions at Gleneagle Secondary; mentor peers in advanced mechanics and observational astronomy.
    \item \textbf{Problem Setter}: Devised competition-level problems for amateur astronomy events and mock Olympiads.
\end{itemize}

\textbf{Science Communication \& Contributions} | \textit{Lead Contributor} \hfill 2022 -- Present
\begin{itemize}
    \item \textbf{FiveHive (Non-Profit aimed at providing help to AP students)}: Authored 18 AP Calculus articles and 7 practice sets for college-prep students.
    \item \textbf{Chinese Wikipedia}: Created 7 professional astronomy articles; significantly improved 12 core scientific entries.
    \item \textbf{Academic Translation}: Voluntarily translated 24 professional astronomy articles (24,700+ words) from English to Chinese.
\end{itemize}

% --- 技术技能 ---
\section{Technical Skills \& Activities}
\begin{itemize}
    \item \textbf{Programming \& Tools}: Python (Data Analysis), Java, Web Development, \LaTeX.
    \item \textbf{Observational Astronomy}: 5 years of professional telescope operation and digital astrophotography.
    \item \textbf{Languages}: English (Fluent), Mandarin (Native), French (Intermediate).
\end{itemize}

\end{document}
